Categorical Predictor Embeddings in Actuarial Modeling

### 2021-03-10 manuscript/manuscript.Rmd
Added Attention and fixed LaTeX

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
extra_dependencies: ["amsmath"]

@@

### Attention {#attention}

>>>>

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
extra_dependencies: ["amsmath", "bm"]

@@

### Attention {#attention}
Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to RNNs [@Bahdanau2015a] and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is now widely used in natural language processing [@Vaswani2017a], and, more recently, has been applied to computer vision and tabular modelling tasks. Attention mechanisms allow deep neural nets to choose covariates entering the model in a flexible manner; these types of applications are demonstrated in, for example, [@Vaswani2017a]. In this section we provide an introduction to attention mechanisms and then discuss applications of attention to modeling tabular data .

An attention mechanism is a mapping:

  ```{=tex}
\begin{equation}
A: \mathbb{R}^{q \times (q \times d) \times (q \times d)} \to \mathbb{R}^q, ~~
  (\mathbf{q}, K, V) ~\mapsto ~ \mathbf{v} = A(\mathbf{q}, K, V),
\label{eq:attention}
\end{equation}
```
where $\mathbf{q} \in \mathbb{R}^q$ is a query vector, $K \in \mathbb{R}^{q \times d}$ is a matrix of key vectors, $V \in \mathbb{R}^{q \times d}$ is a matrix of value vectors and $\mathbf{v} \in \mathbb{R}^q$, the output of the attention mechanism, is a new vector of values. Several different options for the mapping \eqref{eq:attention} exist in the literature; here we first describe the self-attention mechanism used in [@Vaswani2017a]. In more detail, consider that $V$ is a matrix of covariates relating to a regression problem, which can be composed of embeddings of the same dimension for several categorical variables. We wish to apply the attention mechanism to $V$ to estimate the most important of these embeddings for the regression problem. To do this, we formulate the query vector $\mathbf{q}$ which contains information about the regression problem at hand; this is usually done by passing $L$ through a FCN. Similarly, the matrix $K$ contains the relevance of each row of $L$ for the regression problem and is also usually derived by passing each row of $L$ through another FCN. Finally, the matrix $V$ can be set equal to $L$ or alternatively, can be derived as a transformation of $L$ using a third FCN. The first step of the attention mechanism is to derive a vector $\mathbf{a}$ of attention scores:

  ```{=tex}
\begin{equation}
\mathbb{R}^{(1 \times q) \times (q \times d)} \to \mathbb{R}^{1 \times d}, ~~
  (\bold{q}, K) ~\mapsto ~ \mathbf{a} = \langle \mathbf{q}, K^T \mathbb{R}angle.
\label{eq:attention_scores}
\end{equation}
```
The attention scores are then normalized using a soft max function i.e.

```{=tex}
\begin{equation}
a^*_i = \frac{e^{(a_i)}}{\sum_{n=1}^{d} e^{(a_n)}}.
\end{equation}
```
Finally, a linear combination of $V$ is formed:

  ```{=tex}
\begin{equation}
\mathbb{R}^{(1 \times d) \times (d \times q)} \to \mathbb{R}^{1 \times q}, ~~
  (\mathbf{a^*}, V) ~\mapsto ~ \mathbf{v} = \langle \mathbf{a^*}, V \mathbb{R}angle.
\label{eq:attention_output}
\end{equation}
```
